Supplementary Material for ``Evaluating infectious disease forecasts with allocation scoring rules''

### 2023-12-05 alloscore_manuscript/supplement.rnw
some supplement issue edits

For an arbitrary allocation vector $x \in \mathbb{R}^N_{+}$ the expected loss
\begin{align}
\mathbb{E}_{F} [s_A(x, Y)] = \sum_{i=1}^{N} L \cdot \mathbb{E}_{F_i}[(Y_i - x_i)_{+}] \label{eqn:AS_formula}
\end{align} 
is the sum of expected shortages (scaled by $L$) under the allocations $x_i$ in each location. We therefore
have the following necessary condition for $x^{\star} \in \mathbb{R}^N_{+}$ to be an optimal allocation for $\mathbb
{E}_{F} [s_A(x, Y)]$ under the constraint $\sum_{i=1}^{N} x_i = K$: if $\delta > 0$ of the $x_i^{\star}$ units of resource
allocated to location $i$ are reallocated to location $j$, expected shortage will increase in location
$i$ by at least as much as it decreases in location $j$. That is,
\begin{align}
\mathbb{E}_{F_i}[(Y_i - x^{\star}_i - \delta)_{+}] - \mathbb{E}_{F_i}[(Y_i - x^{\star}_i)_{+}] \geq 
\mathbb{E}_{F_j}[(Y_j - x^{\star}_j)_{+}] - \mathbb{E}_{F_j}[(Y_j - x^{\star}_j + \delta)_{+}]. \label{eqn:ASoptimal1}
\end{align}
Since the expected shortages in $i$ and $j$ have right and left derivatives at any $x_i$ and $x_j$ (see Section \ref{sec:shortage}), we can
divide \eqref{eqn:ASoptimal1} by $\delta$ and take limits for $\delta \searrow 0$ to get
\begin{align}
-D_{-} \mathbb{E}_{F} [(Y_i - x^{\star}_i)_{+}] \geq -D_{+}\mathbb{E}_{F} [(Y_j - x^{\star}_j)_{+}]. \label{eqn:RLineq}
\end{align}
Note that the minus signs appear because our optimality condition addresses how a \emph{decrease} in resources will \emph{increase}
the expected shortage in $i$ and vice versa in $j$.
Scaling by $L$ to match the right and left partial derivatives of $\mathbb{E}_{F} [s_A(x, Y)]$ and using formulae \eqref{eqn:dsfl} and \eqref{eqn:dsfr}, \eqref{eqn:RLineq} becomes
\begin{align}
L(1-F_i(x^{\star}_i-)) \geq L(1-F_j(x^{\star}_j)). \label{eqn:RLineq2}
\end{align}
Inequalities \eqref{eqn:RLineq} and \eqref{eqn:RLineq2} remain true with $i$ and $j$ reversed. They hold with $i=j$ as well by the
definition of $F_i(x^{\star}_i-)$. Therefore, a single number $\lambda$ (a \emph{Lagrange multiplier}) exists such that
\begin{align}
L(1-F_i(x^{\star}_i-)) \geq \lambda \geq L(1-F_i(x^{\star}_i)), \quad \text{for all } i \in 1,\ldots,N
\end{align}
that is,
\begin{align}
F_i(x^{\star}_i) \geq 1 - \lambda/L \geq F_i(x^{\star}_i-),
\end{align}
which says (c.f. discussion after \eqref{eqn:F-alpha-ineq} and \eqref{eqn:set-alpha-ineq}) that $x^{\star}_i$ is a quantile $q_{\tau,F_i}$ for
$\tau = 1 - \lambda/L$.

The constraint now determines $\tau$ (and hence the Bayes act) through $\sum_{i=1}^{N} q_{\tau,F_i} = K.$
This equation can be restated as the requirement that 
\begin{align}
K \in TQ_F(\tau) \label{eqn:tau-inclusion}
\end{align}
where the set-valued function
\begin{align}
TQ_F(\tau) := \sum_{i=1}^{N} Q_{F_i}(\tau) = \left[\sum_{i=1}^{N} q^{-}_{\tau,F_i}, \sum_{i=1}^{N} q^{+}_{\tau,F_i}\right]
\end{align}
is defined using interval addition $[a,b] + [c,d] = [a+c, b+d]$.

>>>>

For an arbitrary allocation vector $x \in \mathbb{R}^N_{+}$ the expected loss
\begin{align}
\mathbb{E}_{F} [s_A(x, Y)] = \sum_{i=1}^{N} L \cdot \mathbb{E}_{F_i}[(Y_i - x_i)_{+}] \label{eqn:AS_formula}
\end{align} 
is the sum of expected shortages (scaled by $L$) under the allocations $x_i$ in each location. We therefore
have the following necessary condition for $x^{\star} \in \mathbb{R}^N_{+}$ to be an optimal allocation for $\mathbb
{E}_{F} [s_A(x, Y)]$ under the constraint $\sum_{i=1}^{N} x_i = K$: if $\delta > 0$ of the $x_i^{\star}$ units of resource
allocated to location $i$ are reallocated to location $j$, expected shortage will increase in location
$i$ by at least as much as it decreases in location $j$. That is,
\begin{align}
\mathbb{E}_{F_i}[(Y_i - (x^{\star}_i - \delta))_{+}] - \mathbb{E}_{F_i}[(Y_i - x^{\star}_i)_{+}] \geq 
\mathbb{E}_{F_j}[(Y_j - x^{\star}_j)_{+}] - \mathbb{E}_{F_j}[(Y_j - (x^{\star}_j + \delta))_{+}]. \label{eqn:ASoptimal1}
\end{align}
Since the expected shortages in $i$ and $j$ have right and left derivatives at any $x_i$ and $x_j$ (see Section \ref{sec:shortage}), we can
divide \eqref{eqn:ASoptimal1} by $\delta$ and take limits for $\delta \searrow 0$ to get
\begin{align}
-D_{-} \mathbb{E}_{F} [(Y_i - x^{\star}_i)_{+}] \geq -D_{+}\mathbb{E}_{F} [(Y_j - x^{\star}_j)_{+}]. \label{eqn:RLineq}
\end{align}
Note that the minus signs appear because our optimality condition addresses how a \emph{decrease} in resources will \emph{increase}
the expected shortage in $i$ and vice versa in $j$.
Scaling by $L$ to match the right and left partial derivatives of $\mathbb{E}_{F} [s_A(x, Y)]$ and using formulae \eqref{eqn:dsfl} and \eqref{eqn:dsfr}, \eqref{eqn:RLineq} becomes
\begin{align}
L(1-F_i(x^{\star}_i-)) \geq L(1-F_j(x^{\star}_j)). \label{eqn:RLineq2}
\end{align}
Inequalities \eqref{eqn:RLineq} and \eqref{eqn:RLineq2} remain true with $i$ and $j$ reversed. They hold with $i=j$ as well by the
definition of $F_i(x^{\star}_i-)$. Therefore, a number $\lambda$ (a \emph{Lagrange multiplier}) exists, which is independent of $i$,
such that
\begin{align}
L(1-F_i(x^{\star}_i-)) \geq \lambda \geq L(1-F_i(x^{\star}_i)), \quad \text{for all } i \in 1,\ldots,N.
\end{align}
That is,
\begin{align}
F_i(x^{\star}_i) \geq 1 - \lambda/L \geq F_i(x^{\star}_i-),
\end{align}
which says (c.f. discussion after \eqref{eqn:F-alpha-ineq} and \eqref{eqn:set-alpha-ineq}) that $x^{\star}_i$ is a quantile $q_{\tau,F_i}$ for
$\tau = 1 - \lambda/L$.

The constraint now determines $\tau$ (and hence the Bayes act) through $\sum_{i=1}^{N} q_{\tau,F_i} = K.$
This equation can be restated as the requirement that 
\begin{align}
K \in TQ_F(\tau) \label{eqn:tau-inclusion}
\end{align}
where the set-valued function
\begin{align}
TQ_F(\tau) := \sum_{i=1}^{N} Q_{F_i}(\tau) = \left[\sum_{i=1}^{N} q^{-}_{\tau,F_i}, \sum_{i=1}^{N} q^{+}_{\tau,F_i}\right]
\end{align}
is defined using interval addition $[a,b] + [c,d] = [a+c, b+d]$. (Note that the letter $T$ is being used to connote a ``totalling'' 
operation, as it often is in survey sampling literature.)
